%%%%%%%%%%%%%%%%%%%%%%%%
%% Sample use of the infthesis class to prepare a thesis. This can be used as 
%% a template to produce your own thesis.
%%
%% The title, abstract and so on are taken from Martin Reddy's csthesis class
%% documentation.
%%
%% MEF, October 2002
%%%%%%%%%%%%%%%%%%%%%%%%

%%%%
%% Load the class. Put any options that you want here (see the documentation
%% for the list of options). The following are samples for each type of
%% thesis:
%%
%% Note: you can also specify any of the following options:
%%  logo: put a University of Edinburgh logo onto the title page
%%  frontabs: put the abstract onto the title page
%%  deptreport: produce a title page that fits into a Computer Science
%%      departmental cover [not sure if this actually works]
%%  singlespacing, fullspacing, doublespacing: choose line spacing
%%  oneside, twoside: specify a one-sided or two-sided thesis
%%  10pt, 11pt, 12pt: choose a font size
%%  centrechapter, leftchapter, rightchapter: alignment of chapter headings
%%  sansheadings, normalheadings: headings and captions in sans-serif
%%      (default) or in the same font as the rest of the thesis
%%  [no]listsintoc: put list of figures/tables in table of contents (default:
%%      not)
%%  romanprepages, plainprepages: number the preliminary pages with Roman
%%      numerals (default) or consecutively with the rest of the thesis
%%  parskip: don't indent paragraphs, put a blank line between instead
%%  abbrevs: define a list of useful abbreviations (see documentation)
%%  draft: produce a single-spaced, double-sided thesis with narrow margins
%%
%% For a PhD thesis -- you must also specify a research institute:
%%\documentclass[phd,ilcc,twoside]{infthesis}

%% For an MPhil thesis -- also needs an institute
% \documentclass[mphil,ianc]{infthesis}

%% MSc by Research, which also needs an institute
% \documentclass[mscres,irr]{infthesis}

%% Taught MSc -- specify a particular degree instead. If none is specified,
%% "MSc in Informatics" is used.
% \documentclass[msc,cogsci]{infthesis}
\documentclass[msc, logo]{infthesis}  % for the MSc in Informatics

%% Master of Informatics (5 year degree)
% \documentclass[minf]{infthesis}

%% Undergraduate project -- specify the degree course and project type
%% separately
% \documentclass[bsc]{infthesis}
% \course{Artificial Intelligence and Psychology}
% \project{Fourth Year Project Report}

%% Put any \usepackage commands you want to use right here; the following is 
%% an example:
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{array}
\usepackage{multirow}
\usepackage[table,xcdraw]{xcolor}

%%Package for bullet
\usepackage{enumitem }
%%Package for tick icon in the experiments chapter
\usepackage{pifont}

%%package
\usepackage{listings}

\usepackage{xcolor}
\usepackage{mdframed}

%% This library is used to make 
%% Clickable table of content 
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
 
%% Information about the title, etc.
\title{ Testing SQL-compliance of current DBMSs}
\author{Elias Spanos}

%% If the year of submission is not the current year, uncomment this line and 
%% specify it here:
% \submityear{1785}

%% Optionally, specify the graduation month and year:
% \graduationdate{February 1786}

%% Specify the abstract here.
\abstract{%
Database Management Systems (DBMSs) are widely used in various fields such as the financial sector, the academia and other online services and are very crucial in the day to day management of data. The correct and efficient operation of such systems is an important factor when choosing the right DBMS software. In the past few decades, the amount of data has increased exponentially, causing a rapid increase in the demand for software that can store, organise and manipulate such data. With this requirements in mind, databases and DBMSs such as Oracle DB, Microsoft SQL Server, MySQL and PostgreSQL were created, each implementing its own set of rules. However, since these DBMSs have been implemented with a significant amount of differences, a set of standards was set by the American National Standards Institute (ANSI) in 1986 in order to keep all these implementations to a same standard and ensure compatibility for the SQL language adopted in all implementations. Even though, the SQL Standard has been introduced and adopted by most of the implementations, there are some differences that still exist, probably due to different interpretations of or noncompliance to the aforementioned Standards, and also due to other issues regarding performance of their systems.Since there do exist some differences between each implementation, migrations and changes from one DBMS to another might lead to some bottlenecks such as incompatibility or semantic issues. Therefore, there is a need for alleviating such a problem by exposing such issues between DBMSs, offering proposed solutions that have minimal negative effects to the performance of a system and also for evaluating their conformance to the common Standard. This project will investigate five different implementations and evaluate their conformance to the SQL Standard.  A crucial question that will be investigated by conducting this project is whether DBMSs have been implemented the SQL standards in the same way. The SQL language should pledge that identical SQL code should always return identical answers when it is evaluated on the same database independently of which DBMS is running on. 

The aim of this project is the implementation of a random SQL query generator and a comparison tool for investigating and highlighting the differences that may exist among current DBMSs. Further, we aim to provide a detailed explanation in regards to SQL standards of potential differences and explain how they might affect the transition between current DBMSs. 
 
}

%% Now we start with the actual document.
\begin{document}

%% First, the preliminary pages
\begin{preliminary}

%% This creates the title page
\maketitle

%% Acknowledgements
\begin{acknowledgements}
 
I would like to thank my supervisors Paolo Guagliardo and Leonid Libkin who were always willing to advise me and help me in order to overcome any difficulty. In addition, I want to thank my family who is always by my side 

\end{acknowledgements}

%% Next we need to have the declaration.
\standarddeclaration

%% Finally, a dedication (this is optional -- uncomment the following line if
%% you want one).
% \dedication{To my mummy.}

%% Create the table of contents
\tableofcontents

%% If you want a list of figures or tables, uncomment the appropriate line(s)
%%\listoffigures
%%\listoftables

\end{preliminary}

%%%%%%%%
%% Include your chapter files here. See the sample chapter file for the basic
%% format.

\include{appendix1}
 
\include{Chapters/1-intro}
\include{Chapters/2-background}
\include{Chapters/3-methodology}
\include{Chapters/4-Implementation}
\include{Chapters/5-Expiraments}
\include{Chapters/6-Conclusions}
% \include{chap2}
%% ... etc ...

%%%%%%%%
%% Any appendices should go here. The appendix files should look just like the
%% chapter files.
\appendix
\include{appendix1}
%% ... etc...

%% Choose your favourite bibliography style here.
\bibliographystyle{apalike}

%% If you want the bibliography single-spaced (which is allowed), uncomment
%% the next line.
% \singlespace

\bibliographystyle{abbrv} 

\bibliography{thesis}


%% ... that's all, folks!
\end{document}
